Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2024-08-27 LICS/sections/introduction.tex
added missing github action

Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}. 
This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica_hierarchical_2023}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.

@@

% String diagrams
Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.  There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica_hierarchical_2023} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}. 
% Hierarchical hypergraphs
Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}. 
% String diagram rewrites
Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.

@@

Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.  We overload the binary notation $+$ for our $n$-ary notation.  
We use string diagrams informally in this paper, but it is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.

>>>>

Although our primary interest is theoretical, we believe that phrasing e-graphs in the more general language of category theory will open the door to new potential domains of applications that now use string diagrams and which require the application of optimisation techniques \cite{Selinger_2010, Hasegawa-traced}. 
This may include areas as diverse as digital \cite{ghica_compositional_2023} and quantum circuits
\cite{coecke_interacting_2011,ZX}, functional programs~\cite{ghica-zanassi2023string}, or computational linguistics \cite{wazni_quantum_2022,coecke_lambek_2013}.  
However, in this work we offer little to the e-graph practitioner, as we focus exclusively on foundational aspects, introducing no new applications or improved e-graph algorithmics.

@@

% String diagrams
Our use of string diagrams is also central to the intuition behind our work \cite{Selinger_2010, joyal_geometry_1991}.  There is a breadth of work on similar string diagrammatic formalisms,  including hierarchical string diagrams \cite{ghica-zanassi2023string} and functorial boxes \cite{mellies_functorial_2006},  proof nets for compact closed categories with biproducts \cite{duncan_generalised_2009}, and recent work on tape diagrams for rig categories with biproducts \cite{bonchi_tape_nodate}. 
% Hierarchical hypergraphs
Hierarchical (hyper-)graphs are also a well-studied area of research \cite{plump:hierarchical-graphs, montanari:gs-lambda, palacz:hierarchical-transform, Gaducci:hierarchical-graphs, Ghica:hierarchical}. 
% String diagram rewrites
Our work builds heavily on string diagram rewrite theory as developed in \cite{bonchi_string_2022,bonchi_string_2022-1, bonchi_string_2022-2}.

@@

Figure \ref{fig:egraph-strings} displays the generators of this language and Figure \ref{fig:string-equations} displays the additional equations which these diagrams satisfy, in addition to the standard SMC equations. 
The first four equations are those displayed in Definition \ref{def:enriched-prop},  while the final four axiomatize $+$ as an \textit{n-ary} associative, commutative and idempotent operation.  We overload the binary notation $+$ for our $n$-ary notation.  
It is intuitively clear that these diagrams are sound and complete with respect to their intended categorical semantics. 
Indeed, similar diagrammatic languages using boxes to express choice have been used before~\cite{duncan_generalised_2009}.
